---
title: "Sampling in R"
author: "David"
date: "`r Sys.Date()`"
output: html_document
---

# Probability Sampling

## 1.Simple Random Sampling

In Simple Random Sampling, every individual in the population has an equal chance/probability of being selected to be in the sample. However, the sample may not be representative of the population.

```{r}
# generate a random sample without replacement
sample(1:10,10)
```
```{r}
# Sampling with replacement
sample(1:10,10, replace = TRUE)
```
```{r}
# Including sampling probabilities
# first create a vector of two colors
Colors <- c("green", "red")
Colors
# draw a random sample of 10 colors and green should be more than red
sample(Colors,10, replace = TRUE, prob = c(0.6,0.4))
```

## 2.Stratified Sampling

In stratified sampling, the population is first divided into subgroups based on some characteristic (e.g a nominal variable). The subgroups are called strata. Various samples are then randomly drawn from each stratum, and then they are combined into one whole sample of size n.

```{r}
# using group_by() and sample_n() functions from dplyr

# load the dplyr package
pacman::p_load(dplyr)
# set random seed for reproducibility
set.seed(123)
# group the iris data set by species variable
iris_grouped <- iris |> group_by(Species)
# draw a sample of 3 from each group
sample_n(iris_grouped,3)
```
```{r}
# demonstrate the use of sample_frac() function from dplyr
sample_frac(mtcars,0.2) # selects 20% of the cars
```

```{r}
# stratified sampling using group_by() and slice_sample() functions
mtcars |> group_by(vs) |> slice_sample(n = 4)
```
```{r}
# stratified sampling using strata() function (sampling with replacement)

# first load the sampling package
library(sampling)
# sample
stratas <- strata(iris, c("Species"), size = c(3,3,3), method = "srswr")
stratas # stratified by Species, n1=3, n2=3, n3=3
```
```{r}
# stratified sampling without replacement
New_stratas <- strata(iris, c("Species"), size = c(2,3,4), method = "srswor")
New_stratas # stratified by Species, n1=2, n2=3, n3=4
```

## 3.Systematic Sampling

In Systematic sampling, the members of the population are first arranged in ascending or descending order based on some characteristics. The members of the population are then selected at regular intervals. The starting point (i) is randomly chosen, then k is obtained using the formula k = (N/n). Every kth element is then selected.

```{r}
# first load the "TeachingSampling" package
pacman::p_load(TeachingSampling)

# generate a population to sample from
Population <- c(1:20) # N = 20
Population

# sample, N = 20, k = 5
systematic_sample <- S.SY(20,5)
Population[systematic_sample] # randomly chooses i, k = 5
```

## 4.Cluster Sampling

In Cluster Sampling, the population is first divided into clusters, usually based on geographical areas, and clusters are randomly selected. Each of the selected clusters contain all the members belonging to that particular cluster. One can choose to include all or randomly select members from these clusters.

```{r}
# load the teachers workload data set from the SDaA package
library(SDaA)
data("teachers")

# view the first few observations
head(teachers)

# clean the data
# remove missing values from the data
teachers <- na.omit(teachers)
# list the unique school IDs in the data
unique(teachers[["school"]])
```
```{r}
# Create a cluster sample with 5 randomly selected clusters. (Using school ID as the cluster variable)
cl <- cluster(teachers, clustername = c("school"), size = 5, method = "srswor")
# Sample from the teachers data
cl_data <- getdata(teachers, cl)
# view the first few observations of the sample
head(cl_data)
```
```{r}
# carry out SRS of size 2 from the sampled clusters so that each cluster now contain 2 randomly selected workload details per school cluster.
cl_sample <- cl_data %>% group_by(school) %>% sample_n(size = 2)
# check the unique school IDs for the sampled schools
table(cl_sample$school)

```

Each of the 5 clusters have 2 randomly selected workload data.

```{r}
# view the SRS data from the Clusters
View(cl_sample)
```
